Negative Selection in the 20th Century

### 2020-05-14 why-negative-selection.Rmd
Add income regressions

```{r plot-pc-regs, fig.cap = "Effect of principal components on number of siblings/children. Each dot represents one principal component. Points are jittered on the Y axis.", fig.height = 2}

@@

## Education levels

@@

```{r plot-educ-levels, fig.cap = "Effect sizes on number of children by age left full-time education", fig.height = 7}

loadd(res_edu)
res_edu %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs),
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18")
      ) %>% 
      ggplot(aes(term, estimate, colour = `Age left FTE`)) +
        my_hline +
        geom_pointrange(aes(ymin=conf.low, ymax=conf.high), alpha = 0.8,
          position = position_dodge(width = 0.4)) + 
        scale_colour_viridis_d(begin = 0.1, end = 0.85) +
        coord_flip()
```

>>>>

```{r plot-pc-regs, fig.cap = "Effect of principal components on number of siblings/children. Absolute effect sizes are plotted. Each dot represents one bivariate regression. Points are jittered on the Y axis.", fig.height = 2}

@@

## Education and income

@@

```{r plot-educ-levels, fig.cap = "Effect sizes on number of children by age left full-time education"}

drake::loadd(res_edu)
res_edu %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs),
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18")
      ) %>% 
      ggplot(aes(term, estimate, colour = `Age left FTE`)) +
        my_hline +
        geom_pointrange(aes(ymin=conf.low, ymax=conf.high), alpha = 0.8,
          position = position_dodge(width = 0.4)) + 
        scale_colour_viridis_d(begin = 0.1, end = 0.85) +
        coord_flip()
```

Figure \@ref(fig:plot-income) splits respondents by household income category.
A very similar pattern holds, with selection effects being larger for those
in the poorest income category.

```{r plot-income, fig.cap = "Effect sizes on number of children by household income"}

drake::loadd(res_income)

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs),
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        )
      ) %>% 
      ggplot(aes(term, estimate, fill = Income, colour = p.value < 0.05/{{n_regs}})) +
        my_hline +
        geom_point(alpha = 0.8, shape = 21, size = 1.5, stroke = 0.9) + 
        scale_fill_brewer(direction = -1) +
        scale_colour_manual(values = pval_palette) +
        coord_flip()
```

```{r income-edu-controlled}
drake::loadd(res_income_controlled, res_edu_controlled)
n_regs_ic <- nrow(res_income_controlled)
n_sig_ic <- sum(res_income_controlled$p.value < 0.05/n_regs_ic)

n_regs_educ <- nrow(res_edu_controlled)
n_sig_educ <- sum(res_edu_controlled$p.value < 0.05/n_regs_educ)
```

These results could be driven by age, if older respondents are poorer and less
educated, and also more subject to selection on polygenic scores. However, if we
rerun the regressions, interacting the polygenic score with income category and
also with a quadratic in age, the interaction with income remains significant at
0.05/`r n_regs_ic` for `r n_sig_ic` out of `r n_regs_ic` regressions. Similarly
if we interact the PGS with age of leaving full time education and a quadratic
in age, the interaction with age leaving FTE remains significant at 
0.05/`r n_regs_educ` for `r n_sig_educ` out of `r n_regs_educ` regressions.
